### 2019-09-30 SPsim.Rmd
Simulate one example dataset

---
title: "Simulation study"
output: html_notebook
---

```{r setup}
source('R/SPreg.R', echo=FALSE)
library(ggplot2)
theme_set(theme_classic())
```

Goal:

 * simulate data with a q-power backbone distribution that follows the relative-risk regression model
 * fit the data using our semi-parametric regression model, beta-binomial regression (misspecified correlation structure), and GEE (marginal only)

$$
	\lambda_k(z) = {\mu_k} \times {\theta(\beta' z)^k}, k = 0, \dots, K,
$$

## Simulating data
### Cluster-size distributions

 1. All small (3); 
 2. All large (10); 
 3. varying (0.5*(1+binom(9, 0.5)) + 0.5*Unif(2,10))

```{r Clustersizes}
cs.small <- function(n) rep(3, n)
cs.large <- function(n) rep(10, n)
cs.varied <- function(n) {
  p <- rbinom(n, size=1, prob=0.5)
  x <- ifelse(p, 1+rbinom(n, size=9, prob=0.5), sample(2:10, size=n, replace=TRUE))
  x
}
```

### Covariate-generating functions

Returns model matrix with randomly generated covariates

```{r Covgen}
cov.trt <- function(n, ntrts=4){
  obs <- rep(LETTERS[1:ntrts], length = n)
  dat <- data.frame(Trt = obs)
  dat
}

cov.bin_norm <- function(n, p=0.5){
  bin <- rbinom(n, size=1, prob=p)
  norm <- rnorm(n)
  data.frame(X1 = bin, X2 = norm)
}
```

### Simulate one dataset

The inputs will be

 * `n` - number of clusters to simulate data
 * `clustersize.gen` - function to generate the clustersizes
 * `backbone.gen` - function to generate the PDF of the backbone
 * `covariate.gen` - function to generate the covariates
 * `beta` - coefficients of the linear predictor
 * `link` - the link-function for the semi-parametric model; the inverse link is the $\theta$ function in the equation

```{r SimSetup}
sim.sp <- function(n, clustersize.gen, backbone.gen, 
                   covariate.gen, formula, beta, link="log"){

   model_fun <- binomial(link=link)$linkinv

   cs <- clustersize.gen(n)
   K <- max(cs)
   covs <- covariate.gen(n)
   cov.mat <- model.matrix(formula, data=covs)

   lp <- cov.mat %*% beta
   theta <- model_fun(lp)

   q0 <- backbone.gen(K)
   lambda0 <- lambda_from_p(q0)

   th <- sapply(0:K, function(k)theta^k)
   lambda <- apply(th, 1, function(t)t * lambda0)
   rownames(lambda) <- 0:K

   q <- lapply(1:n,
               function(idx)p_from_lambda(lambda[,idx], n=cs[idx]))
   nresp <- sapply(q, function(qq)sample(seq_along(qq)-1, 
                                         size=1, prob=qq))
   res <- data.frame(ID=1:n,
                     ClusterSize = cs,
                     NResp = nresp)
   res <- cbind(res, covs)
   res
}

```

### Example of simulated data

```{r SimEx}
simd <- sim.sp(n = 100,
               clustersize.gen = cs.varied,
               backbone.gen = function(n)qpower.pdf(p=0.8, rho=0.5, n=n),
               covariate.gen = cov.trt,
               formula = ~Trt,
               beta = c(0, -0.5, -1, -2),
               link = "log")

ggplot(simd, aes(x=Trt, y=NResp/ClusterSize, size=ClusterSize)) +
  geom_jitter(width=0.1, height=0, alpha=0.5) +
  scale_size_area(breaks=c(1,5,10))
```

### Export multiple simulated sets

The GEE and beta-binomial models with log-links will be fitted in SAS.