### 2019-11-22 SPsims/SPsim.Rmd
Move ICOSDA talk simulations to subfolder

---
title: "Simulation study"
output: html_notebook
---

```{r setup}
source('R/SPreg.R', echo=FALSE)
library(knitr)
library(dplyr)
library(tidyr)
library(ggplot2)
theme_set(theme_classic())
```

Goal:

 * simulate data with a q-power backbone distribution that follows the relative-risk regression model
 * fit the data using our semi-parametric regression model, beta-binomial regression (misspecified correlation structure), and GEE (marginal only)

$$
	\lambda_k(z) = {\mu_k} \times {\theta(\beta' z)^k}, k = 0, \dots, K,
$$

## Simulating data
### Cluster-size distributions

 1. All small (3); 
 2. All large (10); 
 3. varying (0.5*(1+binom(9, 0.5)) + 0.5*Unif(2,10))

```{r Clustersizes}
cs.small <- function(n) rep(3, n)
cs.large <- function(n) rep(10, n)
cs.varied <- function(n) {
  p <- rbinom(n, size=1, prob=0.5)
  x <- ifelse(p, 1+rbinom(n, size=9, prob=0.5), sample(2:10, size=n, replace=TRUE))
  x
}
```

### Covariate-generating functions

Returns model matrix with randomly generated covariates

```{r Covgen}
cov.trt <- function(n, ntrts=4){
  obs <- rep(LETTERS[1:ntrts], length = n)
  dat <- data.frame(Trt = obs)
  dat
}

cov.bin_norm <- function(n, p=0.5){
  bin <- rbinom(n, size=1, prob=p)
  norm <- rnorm(n)
  data.frame(X1 = bin, X2 = norm)
}
```

### Simulate one dataset

The inputs will be

 * `n` - number of clusters to simulate data
 * `clustersize.gen` - function to generate the clustersizes
 * `backbone.gen` - function to generate the PDF of the backbone
 * `covariate.gen` - function to generate the covariates
 * `beta` - coefficients of the linear predictor
 * `link` - the link-function for the semi-parametric model; the inverse link is the $\theta$ function in the equation

```{r SimSetup}
sim.sp <- function(n, clustersize.gen, backbone.gen, 
                   covariate.gen, formula, beta, link="log"){

   model_fun <- binomial(link=link)$linkinv

   cs <- clustersize.gen(n)
   K <- max(cs)
   covs <- covariate.gen(n)
   cov.mat <- model.matrix(formula, data=covs)

   lp <- cov.mat %*% beta
   theta <- model_fun(lp)

   q0 <- backbone.gen(K)
   lambda0 <- lambda_from_p(q0)

   th <- sapply(0:K, function(k)theta^k)
   lambda <- apply(th, 1, function(t)t * lambda0)
   rownames(lambda) <- 0:K

   q <- lapply(1:n,
               function(idx)p_from_lambda(lambda[,idx], n=cs[idx]))
   nresp <- sapply(q, function(qq)sample(seq_along(qq)-1, 
                                         size=1, prob=qq))
   res <- data.frame(ID=1:n,
                     ClusterSize = cs,
                     NResp = nresp)
   res <- cbind(res, covs)
   list(data=res, probs=q)
}

```

### Example of simulated data

```{r SimEx}
simd <- sim.sp(n = 100,
               clustersize.gen = cs.varied,
               backbone.gen = function(n)qpower.pdf(p=0.8, rho=0.5, n=n),
               covariate.gen = cov.trt,
               formula = ~Trt,
               beta = c(0, -0.5, -1, -2),
               link = "log")$data

ggplot(simd, aes(x=Trt, y=NResp/ClusterSize, size=ClusterSize)) +
  geom_jitter(width=0.1, height=0, alpha=0.5) +
  scale_size_area(breaks=c(1,5,10))
```

### Simulate and export multiple simulated sets

The GEE and beta-binomial models with log-links will be fitted in SAS, so we will simulate all the datasets at once.

```{r ActualSim}
set.seed(10012019)
R <- 500
sim.list <- 
  lapply(1:R,
         function(idx){
           res <- sim.sp(n = 100,
               clustersize.gen = cs.varied,
               backbone.gen = function(n)qpower.pdf(p=0.5, rho=0.5, n=n),
               covariate.gen = cov.trt,
               formula = ~Trt,
               beta = c(0, -0.5, -1, -2),
               link = "log")$data
           res$Dataset <- idx
           res})
sim.all <- do.call(rbind, sim.list)
write.table(sim.all, file="SimData.txt", row.names=FALSE, quote=FALSE)
```

## Run analyses

```{r RunSAS}
if (.Platform$OS.type == "unix"){
  system("sas FitModels", wait=TRUE)
}
gee.res <- read.delim("GEEest.txt") %>%
  mutate(Coef = paste0(Parm, Level1)) %>%
  select(-Parm,-Level1) %>%
  group_by(Dataset) %>%
  spread(key=Coef, value=Estimate)

bb.res <- read.delim("BBest.txt")%>%
  mutate(Coef = paste0(Effect, Trt)) %>%
  select(-Effect,-Trt) %>%
  group_by(Dataset) %>%
  spread(key=Coef, value=Estimate)
```

```{r RunSP, warning=FALSE}
sp.res.list <- lapply(sim.list,
                 function(d){
                   tryCatch(fit <- sprr(cbind(NResp, ClusterSize-NResp) ~ Trt,
                               data=d, link="log",
                               control=list(maxit=100, eps=0.01)), silent=TRUE,
                            error = function(e)NULL)
                 })
sp.coef <- lapply(sp.res.list, 
                      function(z)if(is.null(z)) rep(NA,4) else coef(z))
sp.mu <- lapply(sp.res.list, 
                 function(z)if(is.null(z)) NA else lambda_from_p(z$q)[2])
sp.res <- as_tibble(do.call(rbind, sp.coef)) %>%
  bind_cols(mu = unlist(sp.mu)) %>%
  mutate(Dataset = 1:length(sp.coef)) %>%
  mutate(Intercept = `(Intercept)`+log(mu))

```

## Compare results

```{r Compare}
comb <- bind_rows(list(GEE=gee.res,
                       BB=bb.res,
                       SP=sp.res), .id="Method") %>%
  select(-TrtA, -`Scale Parameter`, -`(Intercept)`, -mu) %>%
  group_by(Dataset, Method) %>%
  gather(key="Parameter", value="Estimate", Intercept:TrtD)

correct <- tibble(Parameter=c("Intercept","TrtB","TrtC", "TrtD"),
                  Correct = c(log(0.5), -0.5, -1, -2))
comb <- left_join(comb, correct, by="Parameter")

scomb <- comb %>%
  group_by(Method, Parameter) %>%
  summarize(bias = mean(Estimate-Correct, na.rm=TRUE),
            se = sd(Estimate, na.rm=TRUE),
            rmse = sqrt(mean((Estimate-Correct)^2, na.rm=TRUE)))
kable(scomb, digits=2)
```

```{r ComparePlot}
ggplot(comb, aes(x=Method, y=Estimate, colour=Method)) +
  facet_wrap(~Parameter) +
  geom_hline(aes(yintercept=Correct)) +
  geom_jitter(alpha=0.3, height=0)
```

### Compare predicted probability of 0 events in a cluster of size 5

```{r ProbZero}
nd <- expand.grid(Trt=LETTERS[1:4])
sp.zero.list <- lapply(sp.res.list, 
                  function(fit){
                    pr <- predict(fit, newdata=nd, newn=5, type="probvec")
                    cbind(nd, p0=do.call(rbind, pr)[,1])})
sp.zero <- bind_rows(sp.zero.list, .id="Dataset") %>%
  mutate(Dataset = as.numeric(Dataset))

# rho =1/(1+phi)
bb.zero <- bb.res %>% 
  gather(key="TrtS", value="Estimate", TrtA:TrtD) %>%
  mutate(rho = 1/(1+`Scale Parameter`),
         mu = exp(Intercept + Estimate),
         Trt = substr(TrtS, 4, 4),
         p0 = dbetabinom(0, size=5, prob = mu, rho = rho)) %>%
  select(Dataset, Trt, p0)
```

Calculate correct values:
```{r CorrectProbZero}
p0c <- sim.sp(n = 4,
       clustersize.gen = function(n)rep(5,n),
       backbone.gen = function(n)qpower.pdf(p=0.5, rho=0.5, n=n),
       covariate.gen = cov.trt,
       formula = ~Trt,
       beta = c(0, -0.5, -1, -2),
       link = "log")$probs %>%
  sapply(., function(x)x[1])
correct.zero <- cbind(nd, Correct=p0c)
```

```{r ProbZeroComb}
comb.zero <- bind_rows(list(BB=bb.zero,
                            SP=sp.zero), .id="Method") %>%
  left_join(correct.zero, by="Trt")

scomb.zero <- comb.zero %>%
  group_by(Method, Trt) %>%
  summarize(m_p0 = mean(p0, na.rm = TRUE),
            bias = mean(p0 - Correct, na.rm = TRUE),
            se = sd(p0, na.rm=TRUE),
            rmse = sqrt(mean((p0-Correct)^2, na.rm=TRUE)))
scomb.zero
```

```{r CompareZeroPlot}
ggplot(comb.zero, aes(x=Method, y=p0, colour=Method)) +
  facet_wrap(~Trt) +
  geom_hline(aes(yintercept=Correct)) +
  geom_jitter(alpha=0.3, height=0)
```